---
title: "Create Processed Data"
output: html_document
---

*Changes: 1. Replaced ZCTA/MODZCTA API pulls with councilcount's `get_geo_estimates()` call for CD-level estimates (most variables), but kept the CT-level spatial data pull since it is needed to intersect with the park isochrones, and also for the median income variable. 2. Switched from 2010 census (2019 ACS) to 2020 census (2023 ACS, to match councilcount estimates).*

```{r setup, include=FALSE}
source("../code/01_load_dependencies.R") 
```

### Import and Clean ACS Census Tract Data (For Isochrones & Median Income)
```{r ct-data}
# Set census API key
Sys.setenv(CENSUS_API_KEY = "8eb345ebff3eef0f3f10891ef6a297781eefd283")

# Pull 2020 CTs from Open Data 
# Using the geospatial export endpoint bypasses the 1000-row API limit 
url_2020_ct <- "https://data.cityofnewyork.us/api/geospatial/63ge-mke6?method=export&format=GeoJSON"

ct_shp <- sf::read_sf(url_2020_ct) %>%
  # The 2020 file contains an 11-digit 'geoid' (e.g., "36061000100")
  # We append "1400000US" to match the ACS API GEO_ID format exactly
  mutate(GEO_ID = paste0("1400000US", geoid))

# Pull 2023 ACS data at the CT Level (total pop & median income)
ct_pop_inc <- getCensus(
  name = "acs/acs5/subject",
  vintage = 2023, # 2023 ACS to match councilcount vintage
  vars = c("NAME", "GEO_ID", "S0101_C01_001E", "S1901_C01_012E"), 
  region = "tract:*", 
  regionin = "state:36+county:005,047,081,085,061") %>%
  # Clean missing median incomes
  mutate(S1901_C01_012E = ifelse(S1901_C01_012E < 0, NA, S1901_C01_012E))

# Join spatial data with census population & income
ct_acs <- ct_shp %>%
  select(GEO_ID) %>%
  left_join(ct_pop_inc, by = "GEO_ID") %>%
  # Drop geometry-only tracts that have no ACS data (e.g., water bodies, airport runways)
  filter(!is.na(NAME)) %>%
  st_transform("+proj=longlat +datum=WGS84")

st_write(ct_acs, "../data/processed/ct_acs.geojson", driver='GeoJSON', delete_dsn=TRUE)
```

### Pull Council District Demographics via councilcount
```{r cd-data}
# Total Pop: DP02_0088E | Under 18: DP05_0019E | Over 65: DP05_0024E 
# NH White: DP05_0079E | SNAP: DP03_0074E
cd_vars <- c("DP02_0088E", "DP05_0019E", "DP05_0024E", "DP05_0082E", "DP03_0074E")

# Use the Python package to get the 2023 data (returns a standard dataframe)
cd_acs_data <- cc$get_councilcount_estimates(
  acs_year = 2023L, 
  geo = "councildist", 
  var_codes = cd_vars, 
  boundary_year = 2023L
)

# Clean up the Python dataframe column names
cd_acs_data <- cd_acs_data %>%
  rename(
    TotalPop = DP02_0088E,
    Pop_Under18 = DP05_0019E,
    Perc_Under18 = DP05_0019PE,
    Pop_Over65 = DP05_0024E,
    Perc_Over65 = DP05_0024PE,
    Perc_NHWhite = DP05_0082PE,
    Perc_SNAP = DP03_0074PE
  ) %>%
  mutate(councildist = as.character(councildist_2023)) # Ensure matching type for join

# Use the R package to get the 2023 shapes 
cd_shapes <- councilcount::get_geo_estimates(
  acs_year = "2022", 
  geo = "councildist", 
  var_codes = c("DP02_0088E"), # Just pulling a dummy variable to get the shapes
  boundary_year = "2023"
) %>% 
  select(councildist) %>% # Keep ONLY the geometry and the ID
  mutate(councildist = as.character(councildist))

# Join the 2023 Python Data onto the R Map
cd_acs <- cd_shapes %>%
  left_join(cd_acs_data, by = "councildist")

st_write(cd_acs, "../data/processed/cd_acs.geojson", driver='GeoJSON', delete_dsn=TRUE)
```

### Import and Clean Park Maintenance Data 
```{r park-maintenance}
# Updated to 2024 version instead of 2021 that Nick was using
dl <- read_xlsx("../data/raw/AC18-144-Report-DataForWebsite-20241127__67dc7cb4493c8.xlsx") %>%
  clean_names()

pm <- dl %>%
  mutate(property_number = str_extract(property_number, "\\w+"), 
         fixedpost_2024q1 = as.factor(fixedpost_2024q1), 
         fixedpost_2024q2 = as.factor(fixedpost_2024q2), 
         fixedpost_2024q3 = as.factor(fixedpost_2024q3), 
         fixedpost_2024q4 = as.factor(fixedpost_2024q4))

write.csv(pm, "../data/processed/park_maintenance.csv", row.names = FALSE)
```




















# --- Old Code for 2010 Census (only uncomment and run if you want 2019 ACS estimates instead of 2022) ---

### Import and Clean ACS Census Tract Data (For Isochrones & Median Income)
```{r ct-data}
# # Set Census API Key
# Sys.setenv(CENSUS_API_KEY = "8eb345ebff3eef0f3f10891ef6a297781eefd283")
# 
# # 2010 Census Tracts (Clipped to Shoreline)
# url <- "https://www1.nyc.gov/assets/planning/download/zip/data-maps/open-data/nyct2010_22a.zip"
# ct_shp <- sf::read_sf(unzip_sf(url)) %>%
#   mutate(
#     county = case_when(
#       BoroCode == "1" ~ "061", 
#       BoroCode == "2" ~ "005", 
#       BoroCode == "3" ~ "047", 
#       BoroCode == "4" ~ "081", 
#       BoroCode == "5" ~ "085" 
#     ), 
#     GEO_ID = paste0("1400000US36", county, CT2010)
#   )
# 
# # Pull Total Pop (S0101_C01_001E) AND Median Income (S1901_C01_012E)
# ct_pop_inc <- getCensus(
#   name = "acs/acs5/subject",
#   vintage = 2019,
#   vars = c("NAME", "GEO_ID", "S0101_C01_001E", "S1901_C01_012E"), 
#   region = "tract:*", 
#   regionin = "state:36+county:005,047,081,085,061") %>%
#   # The Census uses negative numbers (e.g., -666666666) for missing median incomes. Clean these to NA.
#   mutate(S1901_C01_012E = ifelse(S1901_C01_012E < 0, NA, S1901_C01_012E))
# 
# # Join spatial data with census population & income
# ct_acs <- ct_pop_inc %>%
#   left_join(ct_shp %>% select(!c("county", "BoroCT2010", "BoroCode")), by = "GEO_ID") %>%
#   st_as_sf() %>%
#   st_transform("+proj=longlat +datum=WGS84")
# 
# st_write(ct_acs, "../data/processed/ct_acs.geojson", driver='GeoJSON', delete_dsn=TRUE)
```



# --- Old Code for 2022 ACS (only uncomment and run if you want 2022 ACS estimates instead of 2023, to avoid using Python councilcount) ---

### Pull Council District Demographics via councilcount
```{r}
# Total Pop: DP02_0088E | Under 18: DP05_0019E | Over 65: DP05_0024E 
# NH White: DP05_0079E | SNAP: DP03_0074E
cd_vars <- c("DP02_0088E", "DP05_0019E", "DP05_0024E", "DP05_0079E", "DP03_0074E")

# get_geo_estimates() automatically generates the 'PE' (percentage) columns for us
cd_acs <- get_geo_estimates(acs_year = "2022", geo = "councildist", 
                            var_codes = cd_vars, boundary_year = "2023") %>%
  rename(
    TotalPop = DP02_0088E,
    Pop_Under18 = DP05_0019E,
    Perc_Under18 = DP05_0019PE,
    Pop_Over65 = DP05_0024E,
    Perc_Over65 = DP05_0024PE,
    Perc_NHWhite = DP05_0079PE,
    Perc_SNAP = DP03_0074PE
  )

st_write(cd_acs, "../data/processed/cd_acs.geojson", driver='GeoJSON', delete_dsn=TRUE)
```




